### 2021-08-24 Manuscript/supplemental.Rmd
Worked on manuscript
- Also updated figures: marginalize across intercepts for yield:pArea figures

\newpage

# Appendix A: Supplementary Material {-}

```{r , echo=FALSE, fig.cap='Effect of polygon area and harvest sequence on canola yield. Polygon area is equivalent to ground ', out.width='100%'}  
  include_graphics('../Figures/ModelSummary3b_canola.png',dpi=NA)
```

```{r , echo=FALSE, fig.cap='Temp', out.width='100%'}  
  include_graphics('../Figures/ModelSummary3b_wheat.png',dpi=NA)
```