### 2021-09-03 Manuscript/supplemental.Rmd

\newpage

# Appendix A: Supplementary Material {-}

```{r boundaryTypes, echo=FALSE, fig.cap='Percentage of field boundaries across sampled fields.', out.width='100%'}
bLengths %>% 
  mutate(croptype=datSource$crop) %>% filter(croptype=='Canola'|croptype=='Wheat') %>% 
  # mutate(location=ifelse(grepl('(Alvin|Dean)',filename),'Southern','Central')) %>% 
  select(filename,croptype,contains('prop_')) %>% 
  pivot_longer(contains('prop_')) %>% 
  mutate(name=gsub('prop_','',name),value=value*100) %>% na.omit() %>%
  mutate(name=factor(name,labels=c('Bare','Grassland','Other\nCrop','Shelterbelt','Standard','Wetland'))) %>% 
  ggplot(aes(x=name,y=value,fill=croptype))+geom_boxplot()+
  labs(y='Percentage of field boundary',fill='Crop')+
  scale_fill_manual(values=c('grey50','white'))+
  theme(axis.title.x = element_blank(),legend.position = 'bottom')

# bLengthSummary %>% #Table version
#   kable(.,format='latex',caption='Mean and SD of proportion of boundary types across all fields, along with the number of fields bordered by at least some of the boundary type.',col.names = c('Type','Mean','SD','Number of Fields'),digits = c(1,4,4,1),escape = FALSE)

```

```{r canolaPlotSpeed, echo=FALSE, fig.cap='Effect of ground speed and order (sequence of harvest) on canola yield.', out.width='100%'}  
  include_graphics('../Figures/groundSpeed_canola.png',dpi=NA)
```

```{r wheatPlotSpeed, echo=FALSE, fig.cap='Effect of ground speed and order (sequence of harvest) on wheat yield.', out.width='100%'}  
  include_graphics('../Figures/groundSpeed_wheat.png',dpi=NA)
```