### 2021-09-08 Manuscript/supplemental.Rmd

\newpage

# Appendix A: Supplementary Material {-}

```{r canolaPlotSpeed, echo=FALSE, fig.cap='Effect of ground speed and order (sequence of harvest) on canola yield.', out.width='100%'}  
  include_graphics('../Figures/groundSpeed_canola.png',dpi=NA)
```

```{r wheatPlotSpeed, echo=FALSE, fig.cap='Effect of ground speed and order (sequence of harvest) on wheat yield.', out.width='100%'}  
  include_graphics('../Figures/groundSpeed_wheat.png',dpi=NA)
```

```{r peasPlotSpeed, echo=FALSE, fig.cap='Effect of ground speed and order (sequence of harvest) on pea yield.', out.width='100%'}  
  include_graphics('../Figures/groundSpeed_peas.png',dpi=NA)
```