### 2021-09-09 Manuscript/abstract.Rmd
Added abstract

Field boundaries can improve crop yields by creating better abiotic conditions for crop growth (e.g. mosture, temperature), and can act as refuges for beneficial arthropods.
This suggests that beneficial crop boundaries may create an intermediate hump-shaped increase in crop yield, where negative edge effects are cancelled out by increased ecosystem services (e.g. pollination or pest suppression) from the field boundary.
However, there is little large-scale evidence showing this, largely because yield is costly and time-consuming to measure.
Precision yield data from combine yield monitors has huge potential in this respect, as the equipment is relatively cheap and commonly measured by growers.
<!-- In this study, we used 252 field-years of yield monitor data from three common crops -- wheat (_Triticum aestivum_), canola (_Brassica napus_), or peas (_Pisum sativum_) -- recorded across Alberta, Canada, and examined how yield varied with distances from common crop boundary types. -->
Average crop yield tended to increase with distance from crop boundaries before plateauing at about 50 m, and yield variation (SD) tended to decrease concurrently with distance.
Differences between boundary types tended to be weak and variable, especially for wheat; there was some evidence of certain field boundaries causing a greater decrease in yield at the edge, but these differences tended to be small in magnitude (<5% of yield).
Furthermore, there was little evidence of an intermediate increase in yield, suggesting that a) ecosystem services do not strongly depend on distance or b) their effect is relatively small.
This study represents one of the first uses of precision yield data to measure ecosystem service provision at large spatial scales.